---
title: 'Assignment 2: '
author: "Your name here"
date: "PSTAT 194TR"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```


Perform EDA on  `starwars` dataset which is in the `tidyverse` package. 

Load the package, add appropriate code chunks to answer the questions below as you perform simple EDA techniques on the dataset. Your solution should include both code and your answer in text that interprets the output from the code. 


# `r bfcolor("Exercise 1:", "blue")` EDA: Explore the `starwars` data set

1. Load the package and the  dataset. 

`r bfcolor("Solutions:", "blue")`
```{r}
# add any of your code here
```


2. Take a look at documentation for the dataset. Write two sentences about the dataset.

`r bfcolor("Solutions:", "blue")`
```{r}
# add any of your code here
```


3. What are the dimensions of the data set? How many variables? How many observations?

`r bfcolor("Solutions:", "blue")`
```{r}
# add any of your code here
```

4. What appears to be the unit of observations/samples(rows)? 

`r bfcolor("Solutions:", "blue")`
```{r}
# add any of your code here
```

5. What are the variables(columns)? What are the data types for each column? 

`r bfcolor("Solutions:", "blue")`
```{r}
# add any of your code here
```

6. Pick two variables and calculate two measures of center and two measures of spread. Explain what these measures mean and why you picked these two variables. Optional: Can you calculate these measures in two ways?

`r bfcolor("Solutions:", "blue")`
```{r}
# add any of your code here
```

7. Pick atleast two variables and summarize them visually. Explain what you observe and why you picked these two variables

```{r}
# add any of your code here
```


# Canvas submitting instructions

When you are ready to submit this assignment, 

1. knit to pdf(not html) (Click the grey triangle to right of Knit to see this option)
2. check that everything looks good in the .pdf file
3. Download the pdf file from the server (use Export from the Files pane -> More )
4. Upload your .pdf assignment file to canvas. 

## Optional: 

Create a function to print squares of numbers in sequence.
Call the function supplying 6 as an argument.

```{r}
# Add your code here if necessary
```