---
title: "Work Along : R essentials"
author: "Dr. Uma Ravat"
date: "PSTAT 194TR"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```

# 1. Console vs .Rmd
Compared to working in the R console, .Rmd allows you to rerun (reproduce) code much more effortlessly


`r bfcolor("Exercise 1:", "blue")` **R as a calculator:** use any mathematical operators (+, -, /, and *) to create an expression and make sure it works as expected. _What is `2+2`, `100*3`, `100/10`, `sqrt(25)`?_

_Remember_ You need to add a code chunk and write your expressions in the code chunk 

`r bfcolor("Solutions:", "blue")`




`r bfcolor("Exercise 2:", "blue")` **Object Assignment:** Assign the value of your mathematical operation to the variable `y`. Note the change in the environment pane. 

`r bfcolor("Solutions:", "blue")`




`r bfcolor("Exercise 3:", "blue")` **Load** the `palmerpenguins` **package** so that we have access to all functions and data in this package. 

`r bfcolor("Solutions:", "blue")`




`r bfcolor("Exercise 4:", "blue")` **View** the `penguins` **dataset**


`r bfcolor("Solutions:", "blue")`




`r bfcolor("Exercise 5:", "blue")` Take a look at the `flipper_length` variable


`r bfcolor("Solutions:", "blue")`





`r bfcolor("Exercise 6:", "blue")` **Find the average** `flipper_length`

`r bfcolor("Solutions:", "blue")`




`r bfcolor("Exercise 7:", "blue")` **Get help** on the mean **function**

`r bfcolor("Solutions:", "blue")`




# 2.  Packages, comments, help 


2. Check if the package `nycflights13` is loaded. If not loaded, load it.

3. In a code chunk, any part of the code  starting with '#' is a comment 
4. View the help file for two functions. Pay attention to the sections in the help file. 


