---
title: "Your Turn : R essentials"
author: "Dr. Uma Ravat"
date: "PSTAT 194TR"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```

## 3. R essentials : 

### 3.1 Object assignment, evaluation, errors

Do you understand what each line of code in the code block below is doing? Can you explain each line in your own words to someone?

Notice how the output of code block velow isn't in the knitted ,html file. 

This is because the code block below has an error. Try running it by pressing the run code button(green triangle). 

Read the error to figure out what the issue is. Can you fix the error? What was the error and how did you fix it? 

Once you fix the error, you can change `eval = FALSE` to `eval = TRUE` and the code will get evaluated when knitted and you'll see the code output in the knitted html file. Try it!

```{r eval = FALSE}

x <- 2         # Assignment # note change in environment
x              # Evaluate the expression and print result

y <- 3         # Assignment
y + 5          # Evaluation, y remains 3

z <- x + 17*y  # Assignment # use previously defined object/variable
z              # Evaluation

rm(z)          # Remove z: deletes the object.
z              # Error!

y_cube <- y^3  # use previously defined object/variable y 
#and assign it's cube to the object y_cube

y_cube
y
```

```{r error = TRUE}
y <- 3
Y # R is case sensitive
Y <- 4
Y
y
# Notive how the paste function helps the output be more explicit. 
print(paste("y has value: ", y, ", Y has value:  ", Y))
```

## 5. Data structures

### Vectors

```{r vectors}
# We can create a vector with c(), a function which combines values known as the combine function
v1 <- c(1, 5, 11, 33)       # Numeric vector, length 4
v2 <- c("hello","world")    # Character vector, length 2 (a vector of strings)
v3 <- c(TRUE, TRUE, FALSE)  # Logical vector, length 3, same as c(T, T, F)
```



### Working with vectors

Can you explain how vector operations below work? 

```{r}
# Vector operations: Vectorized operations in R make R efficient!
v1 <- 1:5         # 1,2,3,4,5
v1
v2 <- rep(1,5)    # 1,1,1,1,1 
v2

# Element-wise operations:
v1 + v2      # Element-wise addition
v1 + 1       # Add 1 to each element
v1 * 2       # Multiply each element by 2
v1 + c(1,2)  # Problem: c(1,2) is a vector of a different length; Warning and R still  complete your request by using "recycling"
```
# Extracting specific elements of a vector is called subsetting a vector
```{r}
v1             # Let's recall what v1 is

# Vector elements: Index in R starts at 1
# NOTE: If you are used to languages indexing from 0, R will surprise you by indexing from 1.

v1[3]             # third element of v1
v1[2:4]           # elements 2, 3, 4 of v1
v1[c(1,3)]        # elements 1 and 3 - note that your indexes here are a vector
v1[c(T,T,F,F,F)]  # elements 1 and 2 - only returns the ones that are TRUE
v1[v1>3]          # v1>3 is a logical vector TRUE for elements >3; returns index and not value/element at that index
v1[-3]            # negative number: will return all elements *except* the third
# To replace elements in a vector, assign them a new value
v1[5] <- 10
v1
```