---
title: 'L02: Work Along: Exploring `penguins` dataset in `R`'
author: "Dr. Uma Ravat"
date: "PSTAT 194TR"
subtitle: '<br><br> '
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```

# Simple EDA for `penguins` dataset

```{r}
library(palmerpenguins) # load the package
data(penguins) # load the dataset we want to explore
```

- How many observations/samples? How many variables? 
- What does each observation, variable represent?


##  Data documentation/Data code book


```{r echo=TRUE, eval=TRUE}
?penguins # at console! documentation of `penguins` dataset 
class(penguins) # dataset is stored as a dataframe in R
```


## What kind of data is in `penguins` dataset?

What are the variables(columns), observations/samples(rows), data types?

```{r}
str(penguins)
```

## Viewing data using functions

```{r }
head(penguins)
```
```{r }
tail(penguins)
```

```{r }
dim(penguins)
```

```{r }
ncol(penguins)
```

```{r }
colnames(penguins)
```

```{r }
nrow(penguins)
```

## Descriptive statistics checks in `R`

sample variance = (average) of squared distance from the mean

$$s^2 = \frac{\displaystyle\sum_{i = 1}^{n} (x_i - \bar x )^2}{n-1}$$ 

```{r}
# variance and standard deviation
x <- c(1, 5, 3, 2, 5, 4, 6)
xbar <- mean(x) # mean
x - xbar # distance from mean
1 - xbar; 5 - xbar # check
x[1] - xbar; x[2] - xbar # check
sq_dist <- (x - xbar)^2 # squared distance from the mean
sum(sq_dist) # total(sum) of square distance from the mean
sum(sq_dist)/(length(x)-1)
 # sample variance
var(x) # using R function
sqrt(sum(sq_dist)/(length(x)-1))
sqrt(var(x))
sd(x)
```


```{r}
# quantiles
x <- c(1, 5, 3, 2, 5, 4, 6)
mean(x)
median(x)
quantile(x, .25)
quantile(x, .5)
quantile(x, c(.25, .5, .75))
x
sort(x)

```